---
title: "World Measles Status Dashboard"
output:
  flexdashboard::flex_dashboard:
    theme: readable
    orientation: columns
    source_code: embed
    vertical_layout: fill
runtime: shiny
resource_files:
- data/geo_location.csv
- data/Measle_Incidence.csv
- data/measle_vaccination.csv
---

```{r setup, include=FALSE}
library(flexdashboard)
library(here)
library(shiny)
library(dplyr)
library(tidyr)
library(ggplot2)
```

About
===================================== 

Column {data-width=650}
-----------------------------------------------------------------------

### Purpose
**What is the purpose of this dashboard?**
There exists huge inequality on the measle vaccination coverage and measles related mortality over the world. This descriptive analysis aims to visually depict this inequality, offering the audience a fundamental understanding of the impact of measles vaccination. To effectively address the aforementioned question, I propose dividing it into two distinct parts.

1\. **Global Analysis of Measles Mortality and Vaccination Coverage:**

-   What is the current status of measles incidence rates worldwide?

-   How does vaccination coverage vary across continents?

2\. **Country-Specific Examination:**

-   What is the prevalence and impact of measles in India, Nigeria, China, and the United States?

-   How do vaccination coverage and mortality rates differ among these selected countries?

```{r global}
if (!file.exists("data")) {
  dir.create("data")
}

measles_data <- read.csv(here("data", "Measle_Incidence.csv"))
geo_location <- read.csv(here("data", "geo_location.csv"))
measles_vaccination <- read.csv(here("data", "measle_vaccination.csv"))

measles_data <- measles_data %>% mutate_all(na_if, "")
measles_data <- left_join(measles_data, geo_location, by = c("Country" = "name"))

measles_data <- measles_data %>%
  rename_at(vars(starts_with("x")), ~substr(., 2, nchar(.)))
measles_data <- gather(measles_data, key = "Year", value = "Cases", -Country, -Disease)
measles_data$Cases <- as.integer(gsub(",", "", measles_data$Cases))

vaccination_t <- measles_vaccination %>%
  filter(COVERAGE_CATEGORY == "WUENIC", YEAR %in% c(1980, 1990, 2000, 2010, 2020)) %>%
  select(NAME, YEAR, X1_COVERAGE, X2_COVERAGE)
```

### Data Source

This data is acquired from World Health Organization, which is publicly available and can be downloaded [here](https://immunizationdata.who.int/listing.html?topic=&location=).


Column {data-width=350}
-----------------------------------------------------------------------

### Video Demo

**Here is video overview of dashborad** \n

<iframe width="560" height="315" src="https://www.youtube.com/embed/WFUzPFoXjTs?si=N68oPKWibyupQYhs" title="YouTube video player" frameborder="0" data-external="1" allowfullscreen></iframe>

Data
===================================== 

Column {data-width=650}
-----------------------------------------------------------------------
### **Data Description**

#### Measles Incidence Data
| variable           | class     | description                                          |
|:----------------------|:------------------|:-----------------------------|
| Country            | character | country where Measles disease happen                 |
| Disease            | character | disease name                                         |
| Year               | character | Year of observation                                  |
| Cases              | integer   | number of cases on that observation year             |


#### Measles Vaccine Data
| variable           | class     | description                                          |
|:----------------------|:------------------|:-----------------------------|
| group              | character | level of analysis                                    |
| code               | character | country code                                         |
| name               | character | country name                                         |
| year               | integer   | Year of observation                                  |
| antigen            | character | antigen name                                         |
| antigen_description| character | antigen description                                  |
| coverage_category  | character | data source                                          |
| coverage_category_description| character  | data source description                   |
| taeget_number      | numeric   | target number of coverage                            |
| doses              | numeric   | actual doses                                         |
| x1_coverage        | numeric   | first dose coverage percentage                       |
| x2_coverage        | numeric   | second dose coverage percentage                      |


Column {data-width=350}
-----------------------------------------------------------------------
### **Data Interface**
```{r DTprep, echo=FALSE}
DT_df <- measles_data
DT::renderDataTable({
  DT::datatable(DT_df,
                options = list(autoWidth = TRUE,
                               pageLength = 10,
                               scroller = TRUE,
                               scrollY = '500px'))
})
```


Incidence Rate
===================================== 

In this page, we visualize the 20 countries with highest Measles Incidence in 2022.

```{r}
ui2 <- fluidPage(
  titlePanel("Country Specific Measles Incidence Rate"),
  sidebarLayout(
    sidebarPanel(
      checkboxGroupInput("country", label = "Country",
            choices = c("India" ,"Nigeria" , "Yemen", "Somalia", "Niger",
                        "Ethiopia", "Pakistan", "Liberia", "Indonesia", 
                        "Congo [Republic]", "Zimbabwe", "Afghanistan" , 
                        "Congo [DRC]", "Angola", "Cameroon", "Chad",
                        "Kenya", "Zambia", "Côte d'Ivoire", "Cabon"
                        ),
                        selected = c("India", "Nigeria"))
    ),
    mainPanel(
      plotOutput("bar")
    )
  )
)

server2 <- function(input, output) {
  selected_data <- reactive({
    measles_data %>%
      filter(Year == 2022) %>%
      filter(Country %in% input$country)
  })

  output$bar <- renderPlot({
    ggplot(selected_data(), aes(x= Country, y = Cases, fill = as.factor(Country))) +
      geom_bar(position = "dodge", stat = "identity", color = "royalblue") +
      labs(title = "Country Specific Measles Incidence Rate in 2022",
           x = "Counrty",
           y = "Measles Incidence Rate") +
      theme_grey() + 
      scale_fill_viridis_d(name = "") +
      theme(legend.position="bottom")
  })
}

shinyApp(ui2, server2)
```




Incidence Trend
===================================== 
## Column {.tabset}

### Global Trend
```{r}
global_row <- subset(measles_data, Country == 'Global')
global_row$Year <- as.integer(global_row$Year)
options(scipen = 999)

global_row <- na.omit(global_row)
ggplot(global_row, aes(x = Year, y = Cases, group=1)) +
  geom_line(color = "darkgreen", size = 1.5) +
  labs(title = 'Global Measles Incidence 1980-2022',
       x = 'Year',
       y = 'Global Incidence Rate',
       caption = "Ziying") + 
  theme_grey() + 
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) 
```


### Afghanistan Trend
```{r}
Af_row <- subset(measles_data, Country == 'Afghanistan')
Af_row$Year <- as.integer(Af_row$Year)
options(scipen = 999)

Af_row <- na.omit(Af_row)
ggplot(Af_row, aes(x = Year, y = Cases, group=1)) +
  geom_line(color = "royalblue", size = 1.5) +
  labs(title = 'Afghanistan Incidence 1980-2022',
       x = 'Year',
       y = 'Afghanistan Incidence Rate',
       caption = "Ziying") + 
  theme_grey() + 
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) 
```


### China Trend
```{r}
China_row <- subset(measles_data, Country == 'China')
China_row$Year <- as.integer(China_row$Year)
options(scipen = 999)

China_row <- na.omit(China_row)
ggplot(China_row, aes(x = Year, y = Cases, group=1)) +
  geom_line(color = "red", size = 1.5) +
  labs(title = 'China Measles Incidence 1980-2022',
       x = 'Year',
       y = 'China Incidence Rate',
       caption = "Ziying") + 
  theme_grey() + 
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) 
```


### US Trend
```{r}
US_row <- subset(measles_data, Country == 'United States')
US_row$Year <- as.integer(US_row$Year)
options(scipen = 999)

US_row <- na.omit(US_row)
ggplot(US_row, aes(x = Year, y = Cases, group=1)) +
  geom_line(color = "lightblue", size = 1.5) +
  labs(title = 'US Measles Incidence 1980-2022',
       x = 'Year',
       y = 'US Incidence Rate',
       caption = "Ziying") + 
  theme_grey() + 
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) 
```


Vaccination Coverage Trend
===================================== 

```{r}
summary_data <- vaccination_t %>%
  group_by(YEAR) %>%
  summarize(
    avg_X1_COVERAGE = mean(X1_COVERAGE, na.rm = TRUE),
    min_X1_COVERAGE = min(X1_COVERAGE, na.rm = TRUE),
    max_X1_COVERAGE = max(X1_COVERAGE, na.rm = TRUE),
    avg_X2_COVERAGE = mean(X2_COVERAGE, na.rm = TRUE),
    min_X2_COVERAGE = min(X2_COVERAGE, na.rm = TRUE),
    max_X2_COVERAGE = max(X2_COVERAGE, na.rm = TRUE)
  )

plot_X1 <- ggplot(summary_data, aes(x = YEAR, y = avg_X1_COVERAGE)) +
  geom_point(size=2) +
  geom_line(color = "lightblue", size = 1) +
  geom_errorbar(
    aes(ymin = min_X1_COVERAGE, ymax = max_X1_COVERAGE)) +
  labs(
    x = "Year",
    y = "X1_COVERAGE",
    title = "Average 1-st Dose Over Years with Error Bars",
    caption = "Ziying"
  )  +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size=7),
        axis.title.x = element_text(size = 7, face="bold"),
        axis.title.y = element_text(size = 7, face="bold"),  
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        plot.margin = margin(0.8, 0.8, 0.8, 0.8, "cm"))



# Create a line plot for X2_COVERAGE
plot_X2 <- ggplot(summary_data, aes(x = YEAR, y = avg_X2_COVERAGE)) +
  geom_point(size=2) +
  geom_line(color = "lightpink", size = 1) +
  geom_errorbar(
    aes(ymin = min_X2_COVERAGE, ymax = max_X2_COVERAGE)) +
  labs(
    x = "Year",
    y = "X2_COVERAGE",
    title = "Average 2-nd Dose Coverage Over Years with Error Bars",
    caption = "Ziying"
  ) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size=7),
        axis.title.x = element_text(size = 7, face="bold"),
        axis.title.y = element_text(size = 7, face="bold"),  
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        plot.margin = margin(0.8, 0.8, 0.8, 0.8, "cm")) 

# Show the plots side by side
library(gridExtra)
grid.arrange(plot_X1, plot_X2, ncol = 2)
```


Analysis
===================================== 

Column {data-width=600}
-----------------------------------------------------------------------

**Introduction**

In this session, I will use four countries India, Nigeria, China, and the United States as an example, to show the prevalence and impact of measles in India, Nigeria, China, and the United States. I will also connect the incidence database to the vaccination coverage database to show the relationship between Measles incidence and Measles vaccination in those countries.
```{r}
selected_countries <- c('India', 'Nigeria', 'China', 'United States of America')
Vac_selected <- measles_vaccination %>%
  filter(NAME %in% selected_countries)
```

```{r}
Vac_selected %>%
  ggplot(aes(x = YEAR, y = X1_COVERAGE, group=1)) +
  geom_line() +
  facet_wrap(vars(NAME), scales = "free", ncol = 2) +
  labs(title = "First Dose Measles Vaccination Coverage 1980-2022",
       x = "Year", y = "Measles Vaccination Coverage",
       caption = "Ziying") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```


**Conclusion**

This analysis of vaccine coverage trends reveals a limited advancement, particularly concerning the administration of the second dose. It is imperative for public health scientists to prioritize efforts towards accentuating the critical significance of measles vaccination, with a particular emphasis on the administration of the second dose, underscoring the need for sustained advocacy and intervention in this domain.


Column {data-width=400}
-----------------------------------------------------------------------

**China Specific Analysis**

From the graph above, there is a substantial decline in measles incidence in China between 1980 and 1990. Now, we can specifically investigate the Measles vaccination coverage of China during this period.

```{r}
china_vaccination <- measles_vaccination %>%
  filter(NAME == "China", COVERAGE_CATEGORY == "WUENIC", between(YEAR, 1983, 1993)) %>%
  select(NAME, YEAR, X1_COVERAGE, X2_COVERAGE)
selected_countries <- c('India', 'Nigeria', 'China', 'United States')
Measles_selected <- measles_data %>%
  filter(Country %in% selected_countries)
Measles_selected$Year <- as.integer(Measles_selected$Year)
china_selected <- Measles_selected %>%
  filter(Country == "China", between(Year, 1983, 1993))
china_selected <- left_join(china_selected, china_vaccination, by = c("Year" = "YEAR"))
```

```{r}
china_selected %>%
  ggplot() +
  geom_bar(aes(x = Year, y = Cases), stat = "identity", fill = "lightblue", alpha = 0.5) +
  geom_line(aes(x = Year, y = X1_COVERAGE*5000), color = "darkgreen", size = 1.5)  + 
  labs(
  x = "Year",
  y = "Measles Incidence",
  title = "Incidence and First Vaccination Coverage in China 1983-1993",
  caption = "Ziying"
) +
  scale_y_continuous(sec.axis = sec_axis(~.*0.000001, name="Vaccniation Coverage", labels = scales::percent)) +
  theme_gray() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```


